---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-figs/",
  cache.path = "README-cache/"
)
```

# rrfields

[![Travis-CI Build Status](https://travis-ci.org/seananderson/rrfields.svg?branch=master)](https://travis-ci.org/seananderson/rrfields)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/seananderson/rrfields?branch=master&svg=true)](https://ci.appveyor.com/project/seananderson/rrfields)
<!-- [![codecov](https://codecov.io/github/seananderson/rrfields/branch/master/graphs/badge.svg)](https://codecov.io/github/seananderson/rrfields) -->

The rrfields R package implements spatiotemporal models that allow for extreme
spatial deviations through time. It uses a predictive process approach with
random fields implemented through a multivariate-t distribution instead of a
multivariate normal. The models are fit with [Stan](http://mc-stan.org/).

You can install the development version of the package with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("seananderson/rrfields")
```
    
## An example model

Simulate data:

```{r simulate, cache=TRUE}
library(rrfields)
set.seed(123)
s <- sim_rrfield(df = 3, n_draws = 8, n_knots = 10, gp_scale = 2.5, 
  gp_sigma = 0.2, sd_obs = 0.1)
head(s$dat)
```

```{r plot-sim, cache=TRUE, dependson="simulate"}
print(s$plot)
```

Fit the model:

```{r fit, cache=TRUE, warning=FALSE, message=FALSE, results='hide', dependson="simulate", cache.comments=FALSE}
options(mc.cores = parallel::detectCores()) # for parallel processing
m <- rrfield(y ~ 1, data = s$dat, time = "time",
  lat = "lat", lon = "lon", nknots = 10, iter = 400, chains = 4)
```

```{r print, cache=FALSE, dependson="fit"}
pars <- c("df[1]", "gp_sigma", "sigma[1]", "gp_scale")
print(m, pars = pars)
```

Plot:

```{r plot, cache=TRUE, dependson="fit", warning=FALSE, message=FALSE}
library(bayesplot)
posterior <- rstan::extract(m$model, inc_warmup = FALSE, permuted = FALSE)
mcmc_trace(posterior,  pars = pars)

mm <- as.matrix(m$model)
mcmc_areas(mm, pars = pars[ 1])
mcmc_areas(mm, pars = pars[-1])
```

# References

Predictive-process models:

Latimer, A. M., S. Banerjee, H. Sang Jr, E. S. Mosher, and J. A. Silander Jr. 2009. Hierarchical models facilitate spatial analysis of large data sets: a case study on invasive plant species in the northeastern United States. Ecology Letters 12:144–154.

Shelton, A. O., J. T. Thorson, E. J. Ward, and B. E. Feist. 2014. Spatial semiparametric models improve estimates of species abundance and distribution. Canadian Journal of Fisheries and Aquatic Sciences 71:1655–1666.

...
