---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-figs/",
  cache.path = "README-cache/"
)
```

# rrfields

The rrfields R package implements spatiotemporal models that allow for extreme
spatial deviations through time. It uses a predictive process approach with
random fields implemented through a multivariate-t distribution instead of a
multivariate normal.

You can install the development version of the package with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("seananderson/rrfields")
```
    
## An example model

Simulate data:

```{r simulate, cache=TRUE}
library(rrfields)
set.seed(999)
s <- sim_rrfield(df = 3, n_draws = 15)
```

```{r plot-sim, cache=TRUE, dependson="simulate"}
library(ggplot2)
ggplot(s$dat, aes(x = lon, y = lat, colour = y)) +
  facet_wrap(~time, nrow = 3) +
  geom_point(size = 2) +
  scale_color_gradient2()
```

Fit the model:

```{r fit, cache=TRUE, warning=FALSE, message=FALSE, results='hide', dependson="simulate", cache.comments=FALSE}
options(mc.cores = parallel::detectCores()) # for parallel processing
m <- rrfield(y ~ 1, data = s$dat, time = "time",
  lat = "lat", lon = "lon", nknots = 15, iter = 400, chains = 4)
```

```{r print, cache=TRUE, dependson="fit"}
print(m, pars = c("df[1]", "gp_sigma", "sigma", "gp_scale", "lp__"))
```

Plot:

```{r plot, cache=TRUE, dependson="fit", warning=FALSE}
library(bayesplot)
posterior <- rstan::extract(m, inc_warmup = FALSE, permuted = FALSE)
mcmc_trace(posterior,  pars = c("df[1]", "gp_sigma", "sigma", "gp_scale"))

mm <- as.matrix(m)
mcmc_areas(mm, pars = c("df[1]"))
mcmc_areas(mm, pars = c("gp_sigma", "sigma", "gp_scale"))
```

# References

Predictive-process models:

Latimer, A. M., S. Banerjee, H. Sang Jr, E. S. Mosher, and J. A. Silander Jr. 2009. Hierarchical models facilitate spatial analysis of large data sets: a case study on invasive plant species in the northeastern United States. Ecology Letters 12:144–154.

Shelton, A. O., J. T. Thorson, E. J. Ward, and B. E. Feist. 2014. Spatial semiparametric models improve estimates of species abundance and distribution. Canadian Journal of Fisheries and Aquatic Sciences 71:1655–1666.

...
